SDG contributions

### 2023-09-06 article/article.qmd
add code to reproduce figure 1

```{r load data}
library(sparklyr)
library(ggplot2)
library(patchwork)
library(dplyr)
library(arrow)
library(here)
library(tidyr)
library(forcats)

@@

paper_cols <- c("id", "rank", "doi", "type", "title_normalised", "title", 
                "unknown1", "year", "date", "publisher", paste0("unknown", 2:11),
                "journal", paste0("unknown", c(13, 14)))

@@

papers <- spark_read_csv(sc, "/tklebel/SDG/sdg_papers_collated.csv",
                         name = "papers")
sdg_labels <- spark_read_csv(sc, "/tklebel/SDG/sdg_labels.csv",
                             name = "sdg_labels")
# join the labels here, since we are doing everything by SDG
papers <- papers %>% 
  left_join(sdg_labels)

# remove 2019-2020 altogether. my MAG copy was still incomplete for 2019
papers <- papers %>% 
  filter(year < 2019)
```

@@

>>>>

```{r load data}
library(sparklyr)
library(ggplot2)
library(patchwork)
library(dplyr)
library(arrow)
library(here)
library(tidyr)
library(forcats)
library(readr)

@@

# Overall MAG data
paper_cols <- c("id", "rank", "doi", "type", "title_normalised", "title", 
                "unknown1", "year", "date", "publisher", paste0("unknown", 2:11),
                "journal", paste0("unknown", c(13, 14)))

@@

# SDG specific data
papers <- spark_read_csv(sc, "/tklebel/SDG/sdg_papers_collated.csv",
                         name = "papers")
sdg_labels <- spark_read_csv(sc, "/tklebel/SDG/sdg_labels.csv",
                             name = "sdg_labels")
# join the labels here, since we are doing everything by SDG
papers <- papers %>% 
  left_join(sdg_labels)

# remove 2019-2020 altogether. my MAG copy was still incomplete for 2019
papers <- papers %>% 
  filter(year < 2019)

# Affiliations
affils <- spark_read_csv(sc,
                         "/tklebel/SDG/affiliations_with_country_code.csv",
                         name = "affils")

author_paper_affiliations <- spark_read_csv(
  sc,
  "/tklebel/SDG/sdg_author_paper_affil.csv",
  name = "author_paper_affiliations"
)

# UN country information
un_countries <- read_csv2(here::here("data/external/UNSD — Methodology.csv"))

un_countries_selection <- un_countries %>% 
  select(country_code = `ISO-alpha3 Code`, country_name = `Country or Area`, 
         continent = `Region Name`, sub_continent = `Sub-region Name`)

@@

# Continents over time
```{r}
author_paper_affiliations_w_groups <- make_author_groups(author_paper_affiliations)

papers_w_affils <- papers %>% 
  select(paperid, SDG_label, year, citations_norm) %>% 
  left_join(author_paper_affiliations_w_groups) %>% 
  right_join(affils) %>% 
  select(paperid, SDG_label, year, authorid, affiliationid, country,
         paper_author_cat, author_position, citations_norm)
```

```{r}
papers_per_country_fos_author_pos <- papers_w_affils %>% 
  group_by(country, year, SDG_label, author_position) %>% 
  summarise(n = n()) %>% 
  collect()

papers_per_country_fos_author_pos_country <- papers_per_country_fos_author_pos %>%
  left_join(un_countries_selection, by = c("country" = "country_code")) %>% 
  drop_na()

normalised_region_share_time <- papers_per_country_fos_author_pos_country %>% 
  filter(author_position == "first_author") %>%
  group_by(SDG_label, year, sub_continent) %>% 
  summarise(nn = sum(n)) %>% 
  mutate(share = nn/sum(nn))
```

```{r sub_continents_over_time, fig.width=8.2, fig.height=4.4}
date_scale <- scale_x_date(breaks = as_year(c(2006, 2010, 2015, 2019)),
               date_labels = "%Y")
base_col = "grey80"

subset_countries <- filter(
  normalised_region_share_time, 
  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
                       "Eastern Asia"
  ))

normalised_region_share_time %>% 
  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
  geom_line(colour = base_col) +
  geom_line(data = subset_countries) +
  geom_point(data = subset_countries, size = 1.2) +
  facet_wrap(vars(fix_sdg(SDG_label))) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region within SDGs",
       colour = NULL)
```
